On multilevel constraint decomposition methods \\ for nonlinear variational inequalities

### 2021-07-14 paper/mcd2.tex
Insert a couple of lemmas.

@@

From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.  Note that Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.

The CD methods of Tai \cite{Tai2003} address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.  The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.

@@

We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$ (Algorithm \ref{alg:basiccd}).  (By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.  Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.  If \pr{parallel} $=$ \text{\textbf{true}} then the \text{\textbf{for}} loop can be computed in any order.

\begin{pseudofloat}
\begin{pseudo*}
\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
    if \pr{parallel}: \\+
        for $i \in \{0,\dots,m-1\}$: \\+
            $\hat w_i\in \cK_i$: \\+
                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
    else: \\+
        for $i = 0,\dots,m-1$: \\+
            $\hat w_i\in \cK_i$: \\+
                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
    $\hat w = \sum_i \hat w_i$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}
\caption{The basic constraint decomposition (CD) algorithm for VI problem \eqref{eq:vi}.}
\label{alg:basiccd}
\end{pseudofloat}

The reader may confirm that, inside each boxed VI, the argument of $f$ is an element of $\cK$.  On the other hand, the expression $v_i - \hat w_i$ appears.  To clarify this situation, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
\begin{align*}
[u - R_i u + v_i] - [u - R_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i,  \notag
\end{align*}
for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if
\begin{equation}
e_i = \hat w_i - R_i u \in \cV_i \label{eq:ithupdate}
\end{equation}
is the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.  In particular, note that $u - R_i u + \hat w_i = u + e_i$; the \pr{parallel} version computes the updates $e_i$ independently.

@@

\begin{assumptions*}  \emph{Lipschitz restrictions}: There exists a constant $C_1>0$ so that
\begin{equation}
\left(\sum_{i=0}^{m-1} \|R_i u - R_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
\end{equation}
for all $u,v\in\cK$.  \emph{Totally Lipschitz residual}: There exists $C_2>0$ so that
\begin{equation}
\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:totallylipschitzresidual}
\end{equation}
for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).  Note that the constants $C_1,C_2$ generally depend on $m$.
\end{assumptions*}
% IF PROOF BELOW SUCCEEDS, SO ALGORITHM IS NORM DECREASING, THEN MAY BE ABLE TO ARGUE THAT IT IS O.K. FOR CONSTANTS TO DEPEND ON $u,v,w_{ij}$ etc. AS LONG AS THEY ARE UNIFORMLY BOUNDED ON CERTAIN BOUNDED SETS

@@

>>>>

The following easy, but important, lemma follows directly from $p$-coercivity.  Along the way we define a new quantity $E(v,u^*)$ which is bounded below by a function of the norm error in $v$.  We will use this quantity in the proofs in Section \ref{sec:convergence}.

\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For $v \in \mathcal{K}$,
\begin{equation}
  E(v,u^*) = \ip{f(v)}{v-u^*} - g(v-u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlike}
\end{equation}
\end{lemma}

\begin{proof}
\begin{align*}
E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - g(v-u^*) \\
   &\ge \kappa \|v-u^*\|^p + 0.  \qedhere
\end{align*}
\end{proof}

@@

The CD methods of Tai \cite{Tai2003} address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.  The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.  Note that from now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.  Also, Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.

@@

We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$ (Algorithm \ref{alg:basiccd}).  (By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.  Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.  For the parallel version the \text{\textbf{for}} loop can be computed in any order.

\begin{pseudofloat}[H]
\begin{pseudo*}
\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
    if \pr{parallel}: \\+
        for $i \in \{0,\dots,m-1\}$: \\+
            $\hat w_i\in \cK_i$: \\+
                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
    else: \\+
        for $i = 0,\dots,m-1$: \\+
            $\hat w_i\in \cK_i$: \\+
                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
    $\hat w = \sum_i \hat w_i$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}
\caption{The basic constraint decomposition (CD) algorithm for VI problem \eqref{eq:vi}.}
\label{alg:basiccd}
\end{pseudofloat}

The reader may confirm that, inside each boxed VI, the argument of $f$ is an element of $\cK$.  On the other hand, the expression $v_i - \hat w_i$ appears.  To clarify this situation, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
\begin{align*}
[u - R_i u + v_i] - [u - R_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i,  \notag
\end{align*}
for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if we define
\begin{equation}
e_i = \hat w_i - R_i u \in \cV_i \label{eq:ithupdate}
\end{equation}
as the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.  In particular, note that $u - R_i u + \hat w_i = u + e_i$; the parallel version computes the updates $e_i$ independently.  Also note that for one iteration of the parallel version of the basic CD algorithm, from \eqref{eq:constraintrestrictionsum} and \eqref{eq:ithupdate} observe that
\begin{equation}
\hat w = u + \sum_{i=0}^{m-1} e_i = u^* + \sum_{i=0}^{m-1} \hat w_i - R_i u^*.  \label{eq:ithupdatefacts}
\end{equation}

@@

\begin{assumptions*}  \emph{Totally Lipschitz restrictions}: There exists a constant $C_1>0$ so that
\begin{equation}
\left(\sum_{i=0}^{m-1} \|R_i u - R_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
\end{equation}
for all $u,v\in\cK$.  \emph{Totally Lipschitz residual}: There exists $C_2>0$ so that
\begin{equation}
\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:lipschitzresidual}
\end{equation}
for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).  Note that the constants $C_1,C_2$ generally depend on $m$.
\end{assumptions*}
% IF PROOF BELOW SUCCEEDS, SO ALGORITHM IS NORM DECREASING, THEN MAY BE ABLE TO ARGUE THAT IT IS O.K. FOR CONSTANTS TO DEPEND ON $u,v,w_{ij}$ etc. AS LONG AS THEY ARE UNIFORMLY BOUNDED ON CERTAIN BOUNDED SETS

@@

From the assumptions and the above definitions we have the following core lemma from which a convergence proof for the parallel algorithm is built.

\begin{lemma} \label{lem:core}  Suppose $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  Suppose the constraint decomposition $\mathcal{V}_i,\mathcal{K}_i,R_i$ satisfies \eqref{as:lipschitzrestrictions}, and suppose $f$ satisfies \eqref{as:lipschitzresidual} as well.  Suppose $\hat w$ is computed in the parallel version of Algorithm \ref{alg:basiccd}.  Then
\begin{equation}
   \ip{f(\hat w)}{\hat w - u^*} \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} \|u-u^*\|. \label{eq:core}
\end{equation}
\end{lemma}

\begin{proof}
FIXME: (25) in \cite{Tai2003}
\end{proof}

Note that the coercivity of $f$ is not actually used in proving Lemma \ref{lem:core}.
